---
layout: post
title: "Rural and Evangelical"
date: 2016-12-02
output:
  html_document
share: true
categories: blog
excerpt: "How did rural evangelicals vote in 2016?"
tags: [rstats]
---


```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(foreign)
library(gridExtra)
library(RColorBrewer)
library(choroplethr)
library(choroplethrMaps)
library(viridis)
library(dplyr)
library(DT)
```

I got the county level voting datafile from[here](https://github.com/mkearney/presidential_election_county_results_2016)

I got the 2010 religious census data from [here](http://www.thearda.com/Archive/Files/Descriptions/RCMSCY10.asp)

```{r message=FALSE, warning=FALSE}
vote <- read.csv("D:/2016_election/pres16results.csv", stringsAsFactors = FALSE)
vote$fips <- gsub("(?<![0-9])0+", "", vote$fips, perl = TRUE)
census <- read.dta("D:/2016_election/relcensus.dta", convert.factors = FALSE)
merge <- merge(census, vote, by=c("fips"))
pres <- filter(merge, cand_name == "Donald Trump" | cand_name == "Hillary Clinton")
trump <- filter(merge, cand_name == "Donald Trump")
clinton <- filter(merge, cand_name == "Hillary Clinton")
trump$diff <- trump$pct - clinton$pct
```

## Rural Counties 

There is no set definition of what rural means, so here's what I did: 

```{r message=FALSE, warning=FALSE}
df <- select(trump, cntyname, stabbr, fips, votes, total, pct, POP2010, evanrate)
summary(df$POP2010)
```

I am going to use the counties whose 2010 total population was below the mean of 26,080. Let's see how Trump did in those counties. 


```{r message=FALSE, warning=FALSE}
rural <- filter(df, POP2010 <=26080)
rural$region <- rural$fips
rural$value <- rural$pct
palette_rev <- rev(brewer.pal(8, "RdBu"))
choro = CountyChoropleth$new(rural)
choro$title = "                         How Did Trump Do in Rural Counties?                                "
choro$set_num_colors(1)
choro$ggplot_polygon = geom_polygon(aes(fill = value), color = NA)
choro$ggplot_scale = scale_fill_gradientn(name = "Trump Share", colours = palette_rev)
choro$render()
```

## Counties with Above Average Evangelicals

Let's do the same with evangelicals. 

```{r message=FALSE, warning=FALSE}
summary(df$evanrate)
```

The median here is 190.5 evangelicals per 1000. So any county that's 19.1% evangelical is in this subset. 

```{r message=FALSE, warning=FALSE}
high_evan <- filter(df, evanrate >= 190.5)
high_evan$region <- high_evan$fips
high_evan$value <- high_evan$pct
palette_rev <- rev(brewer.pal(8, "RdBu"))
choro = CountyChoropleth$new(high_evan)
choro$title = "                                   Counties with Above Average Evangelicals"
choro$set_num_colors(1)
choro$ggplot_polygon = geom_polygon(aes(fill = value), color = NA)
choro$ggplot_scale = scale_fill_gradientn(name = "Trump Share", colours = palette_rev)
choro$render()
```

## Let's Combine Rural and High Evangelical

```{r message=FALSE, warning=FALSE}
rural_evan <- filter(df, POP2010 <=26080 & evanrate >= 190.5)
rural_evan$region <- rural_evan$fips
rural_evan$value <- rural_evan$pct
palette_rev <- rev(brewer.pal(8, "RdBu"))
choro = CountyChoropleth$new(rural_evan)
choro$title = "                                   Rural Counties with Above Average Evangelicals"
choro$set_num_colors(1)
choro$ggplot_polygon = geom_polygon(aes(fill = value), color = NA)
choro$ggplot_scale = scale_fill_gradientn(name = "Trump Share", colours = palette_rev)
choro$render()
```

## A Searchable Table

```{r message=FALSE, warning=FALSE}
table <- select(rural_evan, cntyname, stabbr, pct, POP2010, evanrate)
table$pct <- round(table$pct, 2)
table$evanrate <- round(table$evanrate, 2)
datatable(table, colnames =  c("County", "State", "Trump's Percentage", "2010 Population", "Total Evangelicals per 1000"))
```

## So, how many rural, highly evangelical counties voted for Hillary Clinton?

```{r message=FALSE, warning=FALSE}
dim(rural_evan)
hrc <- filter(rural_evan, pct <=.5)
dim(hrc)

```

There were 852 counties in the total dataset and just 63 had a majority of the votes for Clinton. Trump carried 92.6% of those counties.  


